Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach

### 2018-08-16 paper/competing_bandits.tex
finally finished a draft

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{natbib}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{float}
\usepackage{sgame, tikz} % Game theory packages
\usepackage{caption} 
\usepackage{algorithm,algpseudocode}
\usepackage{makecell}
 \usepackage{multirow}
 \usepackage{graphicx}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\captionsetup{font=footnotesize}
\bibliographystyle{aaai}

\title{Competing Bandits}

@@

\author{-}
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\

@@

\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}
\label{S:1}

...

\section{Related Work}
\label{S:2}

\citealt*{mansour2017competing}, \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}, etc.

\section{Model}
\label{S:3}

...

\section{Simulation Details}
\label{S:4}

\subsection{Bandit Priors}
We study the implications of our model via simulation. For computational tractability we focus our attention on Bernoulli bandit diBernoulli bandits with beta priors. We consider the following bandit ``priors" from which the bandit instances faced in simulation are drawn:
\begin{enumerate}
\item ``HeavyTail": We took the mean rewards to be randomly drawn from Beta($\alpha=0.6,\beta=0.6$). With this distribution it was likely to have arms that were at the extremes (close to 1 and close to 0) but also some of the arms with intermediate value means. [We should rename this prior instance]
\item Needle-in-haystack High - $K-1$ arms with mean 0.50, 1 arm with mean 0.70. 
\end{enumerate}

[Need some discussion of what makes these priors at all ``representative" - why can we just focus on them?]

[We can't feasibly include all the priors. I think Heavy Tail is interesting because of the relative reputation plots, but I'm not sure what other one to include that would be meaningful. For now I've just used NIH]

In the appendix we consider the performance of additional priors, but qualitatively the results presented are unchanged.

\subsection{Learning Algorithms}

For our simulations we evaluate the performance of three algorithms, each endowed with a ``fake prior" of $Beta(1, 1)$. In general, there are three different classes of learning algorithms and we select one representative algorithm from each class: 
\begin{enumerate}
\item Adaptive exploration algorithms that engage in purposeful exploration in a ``smart" way by adapting to the previous history of results. We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward.
\item Non-adaptive exploration algorithms that engage in purposeful exploration without considering the past results. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / Myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}

Run in isolation we expect that the performance of the algorithms (according to cumulative regret) should be \\
 Adaptive Exploration $\geq$ Non-adaptive exploration $\geq$ Greedy. We verify that, for the chosen algorithms and chosen priors, this is indeed the case (with some caveats, discussed in the next section).

The primary question that we are interested in is under what conditions in competition are the firms in our model incentivized to commit to adaptive exploration algorithms.

\subsection{Simulation of Competition Game}
Algorithm \ref{alg_1} (FIX TO BE FIGURE) describes the simulation procedure utilized to evaluate the results of the competition game. $T$ is the finite time horizon of the competition game and $K$ is the number of arms in the bandit instance that is considered.

\begin{algorithm}
\caption{Simulation Pseudo-Code}
\begin{algorithmic}[1]
\For{Each prior $p$}
	\For{$N$ simulations}
		\State Generate true distribution from $p$
		\State Generate $T \times K$ realizations for the arms 
		\For{Each agent algorithm $agent alg$}
			\For{Each principal algorithm pair $principalalg1$, $principalalg2$}
				\State Give principal 2 $X$ free observations (the agents also get these observations)
				\State Give the agents $k$ observations from each principal (warm start)
				\State Run simulation for $T$ periods
			\EndFor
		\EndFor
	\EndFor
\EndFor
\end{algorithmic}
\label{alg_1}
\end{algorithm}

\section{Performance in Isolation}
\label{S:5}

We look at the performance of each of our learning algorithms in isolation, meaning that we simply run the learning algorithm for $T$ steps on a realization of the prior instances that we consider and record the realized rewards and the arm that is pulled. In our simulations the realizations of each of the arms were pre-drawn and the same across all algorithms and so the differences in algorithm performance come from the difference in the arms selected by the algorithms and not from differences in realizations. When looking at the results of each of the learning algorithms in isolation we want to confirm that, for the prior instances that we consider, ``better" algorithms win according to ``absolute" measures of performance (i.e. looking at mean reputation). Additionally, we want to understand what statistics we can look at in isolation to understand what wins under competition.

In our model the reputation score is an input to the decision rule of the agents and thus it is important to understand how this score differs across algorithms when run outside of competition. Figure \ref{prelim_means} displays the mean average reputation score of each of the algorithms across the different simulations.

\begin{figure}
\caption{Mean Reputation Trajectories in Isolation}
\includegraphics[scale=0.2]{"figures/heavy_tail_iso_mean"}
\includegraphics[scale=0.2]{"figures/nih_iso_mean"}
\label{prelim_means}
\caption*{\tiny{The plots contain the average reputation over $250$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs. TODO: get rid of smoothing on plot}}
\end{figure}

\begin{figure}
\caption{Relative Reputation Plots}
\includegraphics[scale=0.2]{"figures/deg_dg_ht_10_prelim"}
\includegraphics[scale=0.2]{"figures/ts_dg_ht_10_prelim"} \\
\includegraphics[scale=0.2]{"figures/deg_dg_nih_10_prelim"}
\includegraphics[scale=0.2]{"figures/ts_dg_nih_10_prelim"} \\
\caption*{\tiny{The plots contain the proportion of simulations over $250$ runs for a memory size of $100$ where, for a given $t$, the reputation of one algorithm is at least as high as the reputation of another on the same instance.}}
\label{relative_rep}
\end{figure}

As expected we see that according to the mean reputation plots $TS > DEG > DG$ for sufficiently large values of $t$. While looking at the mean reputation is a natural statistic to consider when attempting to understand what algorithm wins under competition, we consider an additional statistic when attempting to understand what wins under competition. In the competition game the decision rule of the agents compares the relative reputation of the two firms so that if the reputation of firm $A$ is higher than that of firm $B$, the agents will (be more likely to) select firm $A$. Thus, another natural statistic to look at is to look at the two algorithms in isolation and consider the fraction of simulations where, for a fixed time $t$, the reputation of firm $A$ is at least as high as that of firm $B$. Figure \ref{prelim_comparison} displays the proportion of rounds where one algorithm has at least as high of a reputation as another.

Interestingly we observe that, on the Heavy Tail prior, $DG > DEG$ according to this metric even though $DEG > DG$ according to the mean reputation. The intuition for this is simple. $DG$ will either find the best arm or not. If it does then since $DEG$ engages in non-adaptive exploration $DG$ will, with $\epsilon$ probability, pick an arm uniformly at random which potentially hurts its reputation if the selected arm is not the best arm. $DEG$ eventually converges on the best arm with a sufficiently large time horizon, but $DG$ does not necessarily do so and so when $DG$ does not find the best arm it may potentially have substantially worse reputation which drags down its mean reputation. However, if it is able to find the best arm in sufficiently many simulations then it may have slightly higher reputation than $DEG$ in those simulations and thus do better according to the relative reputation proportion but, due to the simulations where it does not find the best arm it may then have worse mean reputation. In the next section we move to considering the competition game and revisit if this statistic is more useful for predicting who wins in competition.

\section{Simultaneous Entry Experiment}
\label{S:6}

In this section we present simulation results of the model discussed in Section \ref{S:2} where $X = 0$ so that both firms enter at the same time. We want to understand what algorithms the firms are incentivized to commit to in the competition game but the payoffs associated with each algorithm are too difficult to compute analytically and thus, we begin by simply imposing that the algorithm played by the firm is taken as given and study the consequences of playing them in the competition game. Additionally, for this experiment we present results for the $HardMax$ agent decision rule and postpone discussion of alternative decision rules. Besides keeping track of market share, we also will keep track of the ``effective end of game."

\begin{definition}
\textit{Effective End of Game (EEOG)} - the last round, $t$, in a simulation where there is a difference between the decision of the agent alive at $t-1$ and the agent alive at $t$ (todo: state this formally using notation of model)
\end{definition}

\begin{table}[ht]
\centering
\caption{Simultaneous Entry Needle In Haystack} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 100 & k = 200 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.64} $\pm$0.03\\ eeog \\ avg: 210\\ med: 28} & \makecell{\textbf{0.62} $\pm$0.03\\ eeog \\ avg: 260\\ med: 68} & \makecell{\textbf{0.6} $\pm$0.03\\ eeog \\ avg: 320\\ med: 0} \\ 
  TS vs DEG & \makecell{\textbf{0.56} $\pm$0.03\\ eeog \\ avg: 150\\ med: 11} & \makecell{\textbf{0.54} $\pm$0.03\\ eeog \\ avg: 240\\ med: 24.5} & \makecell{\textbf{0.51} $\pm$0.03\\ eeog \\ avg: 420\\ med: 36.5} \\ 
  DG vs DEG & \makecell{\textbf{0.43} $\pm$0.03\\ eeog \\ avg: 330\\ med: 97.5} & \makecell{\textbf{0.42} $\pm$0.03\\ eeog \\ avg: 550\\ med: 229.5} & \makecell{\textbf{0.41} $\pm$0.02\\ eeog \\ avg: 650\\ med: 420} \\ 
   \hline
   \label{hm_nih}
\end{tabular}
\caption*{\tiny{The first line in each cell contains the average market share received by the firm playing Alg 1 (and the market share of Alg 2 is 1 - Alg 1 Market Share) as well as a 95 \% confidence band. For example, the cell in the top left indicates that TS gets on average 32\% of the market when played against DG. The next line contain the average and median effective end of game for this set of simulations.}}
\end{table}

\begin{table}[ht]
\centering
\caption{Simultaneous Entry Heavy Tail} 
\begin{tabular}{rlll}
  \hline
 & k = 20 & k = 100 & k = 200 \\ 
  \hline
TS vs DG & \makecell{\textbf{0.29} $\pm$0.03\\ eeog \\ avg: 65\\ med: 0} & \makecell{\textbf{0.52} $\pm$0.03\\ eeog \\ avg: 130\\ med: 0} & \makecell{\textbf{0.72} $\pm$0.02\\ eeog \\ avg: 500\\ med: 0} \\ 
  TS vs DEG & \makecell{\textbf{0.3} $\pm$0.03\\ eeog \\ avg: 26\\ med: 0} & \makecell{\textbf{0.59} $\pm$0.03\\ eeog \\ avg: 160\\ med: 0} & \makecell{\textbf{0.86} $\pm$0.02\\ eeog \\ avg: 420\\ med: 0} \\ 
  DG vs DEG & \makecell{\textbf{0.63} $\pm$0.03\\ eeog \\ avg: 390\\ med: 10} & \makecell{\textbf{0.62} $\pm$0.02\\ eeog \\ avg: 670\\ med: 500.5} & \makecell{\textbf{0.61} $\pm$0.02\\ eeog \\ avg: 770\\ med: 670.5} \\ 
   \hline
   \label{hm_ht}
\end{tabular}
\end{table}

[IDEA: take a look at the distribution of reputation induced by each algorithm under each prior]

Tables \ref{hm_nih} and \ref{hm_ht} display the results. Under Heavy Tail we see that for small warm start of $k = 20$, $DG$ gives the firm more of the market than $TS$ and $DEG$. As we increase the warm start to $k = 100, 200$ we observe that $TS$ does better than $DG$ and $DEG$. Note that, in particular for the simulations involving $TS$, the effective end of game had a median of 0. This means that the result of the simulation was \textit{completely determined by the warm start period}. Thus, when reasoning about what algorithm would win in the competition game under the HardMax decision rule we need to think about the performance of the algorithm on $k$ samples instead of $T$ samples, which may be small. For low warm start, it is thus not necessarily the case that $TS$ is the best algorithm. However, we see that under the Needle In Haystack prior $TS$ is always the best algorithm across all warm start values.

[Missing here is a discussion of the fact that with HardMax the market gets taken by one firm completely (in most simulations) as well as why this is the case.]

First, we note that the relative reputation plots were better at predicting the result of the competition game than the ``absolute" mean reputation plots. For the Needle In Haystack prior, both plots showed that $TS > DEG > DG$ in isolation and this is also the result of the competition game. However, for the Heavy Tail prior, we see that $DG, DEG > TS$ for low warm start and that, in general, $DEG > DG$ \footnote{The result that $DEG > DG$ for this prior is robust across decision rules as well.}. While the ``absolute" reputation plots imply that $TS > DEG > DG$, we see that the relative reputation plot in Figure \ref{relative_rep} implies that $DEG > DG$ and in Figure \ref{rel_rep_ht_fine} we can see that in the early rounds, $DG$ does relatively better than $TS$ and that, around $t = 100$, $TS$ does relatively better than $DG$. Note that since the median effective end of game is $0$ in \ref{hm_ht} and the agents are using the $HardMax$ decision rule, this means that the relative reputation between the two firms after $k$ samples ends up determining who takes the market in the simulations and this is precisely what the relative reputation plots tell us. Thus, looking at the relative reputation plot at time $k$ gives us a good predictor of which algorithm will take more of the market in the competition game.

Why is it that we see that $TS$ always wins under Needle in Haystack but, for lower warm start, $DG > TS$? We know that for a sufficiently long time horizon, $TS$ leads to a higher reputation than $DG$. However, in our model, firms need to incentivize the agents to select them over their competition. One possible intuition for the results here are that, in the early rounds, $TS$ engages in purposeful exploration whereas $DG$ is myopic. A consequence of exploration in our model is that it may lead to a reputational cost relative to the greedy algorithm since the sub-optimal arms pulled during exploration can harm reputation relative to the greedy algorithm. $TS$ thus incurs a relative reputational cost in the early rounds, but once it has explored enough and gained enough information it is able to recover the relative reputational loss and acquire a larger reputation than the greedy algorithm. A sufficiently large warm start allows $TS$ to explore in the early rounds and recover the reputational loss and thus win the competition game whereas a low warm start is not sufficient for this and thus $DG$ beats $TS$. Why do we only see this on Heavy Tail and not Needle in Haystack? One possible explanation for this is that learning on Heavy Tail is harder (use learning hardness metric here) and learning on Needle In Haystack is easier and thus TS needs more samples in the Heavy Tail instances compared to the Needle In Haystack instances in order to be able to recover the relative reputation loss [TODO: Explain this a bit more and with more formality. Can we have some experimental results that ``prove" this or generalize it?]

[TODO: Can we say anything general about this using the ``learning hardness" metric that we have started looking at?]

\section{Incumbent Experiment}
\label{S:7}

We now fix the same parameters as before, fixing a warm start of $k = 20$ [we should re-run this] but let $X = 200$ so that firm 2 has an ``incumbency" advantage. Recall that $X$ controls the number of rounds the firm is in the market by itself before the other firm enters and so in these rounds agents arrive and must choose firm 2. Importantly, we treat $X$ as an exogenous parameter (and not something optimally chosen by the entrant) and study the consequences of a fixed value of $X$. Thus, firm 2 is a monopolist for $X$ periods and, after $X$ rounds, firm 1 enters the market and each agent that enters in subsequent periods must select between the two firms. We want to understand what algorithms the incumbent (and the entrant) are incentivized to play. In particular, we are interested if introducing this asymmetry in entry induces firms who were previously incentivized to play $DG$ to play $TS$. As a result, we present the results of this experiment for $k = 20$ and on the Heavy Tail prior, where we saw that $DG$ was preferred to $TS$.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amssymb}
\usepackage{natbib}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{float}
\usepackage{sgame, tikz} % Game theory packages
\usepackage{caption} 
\usepackage{algorithm,algpseudocode}
\usepackage{makecell}
 \usepackage{multirow}
 \usepackage{graphicx}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\captionsetup{font=footnotesize}
\bibliographystyle{aaai}

@@

@@

\title{Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach}

\author{Guy Aridor\textsuperscript{1}, Kevin Liu\textsuperscript{2}, Aleksandrs Slivkins\textsuperscript{3},
Zhiwei Steven Wu\textsuperscript{4} \\,
{\textsuperscript{1}Columbia University, Department of Economics}\\
{\textsuperscript{2}Columbia University, Department of Computer Science}\\
{\textsuperscript{3}Microsoft Research, New York, NY}\\
{\textsuperscript{4}University of Minnesota - Twin Cities, Department of Computer Science}
}
\maketitle

\begin{abstract}
We empirically study the interplay between \textit{exploration} and \textit{competition} - how systems that learn from interactions with users tradeoff between making potentially suboptimal choices in order to acquire new information and the reputational consequences of this exploration which potentially has adverse competitive effects. Our model considers competition for myopic users between two firms deploying multi-armed bandit algorithms that face the same underlying multi-armed bandit instance. The agents select between the firms according to a reputation score, which is a function of the rewards past agents have experienced from this firm.

We ask whether better algorithms are incentivized under varying degrees of competition. The environments we consider are a monopoly, a duopoly with one firm serving as an early entrant (or a temporary monopoly), and a duopoly. We show that, under certain conditions, monopoly and duopoly do not incentivize the adoption of better learning algorithms due to the reputational costs of exploration, but that a temporary monopoly may incentivize the adoption of better learning algorithms. Finally, we ask whether in this setting the data advantage or the reputational advantage of the early entrant serves as stronger barriers to entry.
\end{abstract}

\section{Introduction}
\label{intro}

Many modern online platforms simultaneously compete for users as well as learn from the users they manage to attract. For instance, Google Search and Bing compete for users in the search engine market yet at the same need to experiment with their search algorithms to learn what algorithms work the best. This creates a tradeoff between \textit{exploration} and \textit{competition} since firms need to experiment with potentially sub-optimal options for the benefit of gaining information to make better decisions tomorrow while at the same time firms need to incentivize consumers to select them over their competitors today.

From a social welfare perspective it is better for firms to adopt ``smarter" learning algorithms and our main high-level question is asking under what levels of competition are firms incentivized to adopt ``better" learning algorithms. Our primary contribution is that in our model, even in the case of a duopoly, firms are not incentivized to adopt better learning algorithms but that allowing one firm to be a temporary monopolist will incentivize this firm to play a better learning algorithm.

Unlike in the classic models in economics regarding the interplay between competition and innovation described in \citet{barro2004economic}, here we have no explicit R \& D cost. Rather, firms are not incentivized to experiment under duopoly because experimentation has an implicit relative \textit{reputational} cost compared to the myopic alternative and, in our model, consumers only care about the reputation of a firm. Thus, competition disincentivizes experimentation by firms precisely because competition forces firms to take actions in order to incentivize consumers to select them over their competitors and, as a result, by allowing a firm to be a temporary monopolist the firm can experiment without worrying about the reputational consequences. Our findings give an alternative intuition to the empirically documented inverse-U relationship between competition and innovation discussed in \citet{aghion2005competition}.

\textbf{Related Work} A major underlying component of our model is that firms face a multi-armed bandit (MAB) problem. Multi-armed bandits (MAB) are a tractable abstraction for the tradeoff between exploration and exploitation. MAB problems have been studied for many decades (see \citet{bubeck2012regret} for an overview). Most relevant to this paper is the thread that focuses on designing ``smart" and tractable algorithms that combine exploration and exploitation and ``naive" algorithms that separate exploration and exploitation (see \citet{slivkins2017bandits}).

The three-way tradeoff between exploration, exploitation, and incentives has been studied in several other settings, the most relevant to this paper being \citealt*{che2017recommender}, \citealt*{kremer2014implementing}, \citealt*{mansour2015bayesian}. The strategic experimentation literature in economics, such as \citealt*{bolton1999strategic} and \citealt*{keller2005strategic}, studies models with self-interested agents jointly performing exploration whereas in this paper the firms cannot observe the actions or the payoffs of the other firms and exploration is coordinated by the consumers.

The relationship between competition and innovation has been heavily studied in industrial organization \citep{tirole1988theory} and endogenous growth theory \citep{aghion2005competition, barro2004economic}, dating back to \citet{schumpeter2010capitalism}.

Our setting is also closely related to the ``dueling algorithms" framework described in \cite{immorlica2011dueling}, but this framework considers offline scenarios whereas we focus on online learning problems.

The most closely related work to this paper is \citealt*{mansour2017competing}. Our motivating questions are the same as \citet{mansour2017competing} though while \citet{mansour2017competing} use the rationality of consumers as their primary knob of competitiveness, we consider differences in the number of firms in the market as our primary knob. Additionally, while in \citet{mansour2017competing} agents select firms based on the Bayesian expected reward of each principal in our model the agents select firms based on a frequentist reputation score that is derived from the signals sent from past agents. This allows the model to become tractable for the purposes of evaluating it via simulation and this is the primary method of analysis that we employ.

\section{Model}
\label{model}

\textbf{Overview} There are two principals and $T+2k$ agents where $k$ is the warm start, or the agents that each principal gets for free at the beginning of the game and $T$ is the number of rounds in the game. The timing of events is as follows:
\begin{enumerate}
\item At $t = 0$, the principals simultaneously commit to following a learning algorithm from a set of algorithms $\mathcal{A}$
\item Still at $t=0$ we suppose that each principal gets $k$ agents as a ``warm" start.
\item A new agent arrives each round (and lives for only one round), starting at $t = 1$, and chooses among 2 principals, given a reputation score for both principals.
\item The principal that is chosen selects an action $a_{t} \in A$ from a set of fixed actions across principals and rounds.
\item Both the agent and the principal observe the reward $r_t \in [0, 1]$ from the action. The agent reports this reward to the principal and the future agents and the reputation score for the chosen principal is updated.
\item Repeat 2-4 for $T$ rounds.
\end{enumerate}

Generally, the rewards are i.i.d. with a common prior.  For computational tractability we restrict our focus to Bernoulli-distributed rewards with Beta priors. Each principal faces a multi-armed bandit problem with no initial information \footnote{For algorithms that require some sort of prior to operate, such as Thompson Sampling, we use a ``fake" prior of $Beta(\alpha=1,\beta=1)$}. We assume that the principals commit to a multi-armed bandit learning algorithm at the start of the world and that there are no informational spillovers from their competitors so that they can only learn from the agents that select them.

\noindent \textbf{Agents} We suppose that agents are homogenous, myopic, and non-strategic. The utility function for the agents is simply to maximize their reward in the one period in which they are alive. We suppose that agents do not attempt to manipulate the strategy of the principal nor do they take the strategies of the principal into account when choosing between the principals. In our model, each agent uses the average reward of past agents as a proxy for their expected utility. For simplicity, the reputation score, $R_jt$ is defined as a sliding window average.
\begin{center}
$R_{jt} = \frac{1}{M} \sum\limits_{i=1}^{M} r_{t_j-i}$
\end{center}

Note that the reputation score is in terms of \textit{local} and not global time so that it is the sliding window average of the last $M$ times that principal $j$ was selected by the agents. The warm start of $k$ rounds allows this reputation score to be well-defined once the ``competition game" begins. In our model the agents deterministically choose the principal with the higher reputation score and ties are broken uniformly at random.

\noindent \textbf{Principals} We suppose that the principals simply care about maximizing their expected market share (i.e. maximizing the number of $T$ agents who select them). We model the ``competition" game between principals as a simultaneous move game where both principals commit to a learning algorithm.

\noindent \textbf{MAB algorithms} We suppose that principals commit to a learning algorithm from $\mathcal{A}$. We partition the set of possible learning algorithms into three different types of learning algorithms and restrict $\mathcal{A}$ to contain a representative each algorithm from each class:
\begin{enumerate}
\item ``Smart" algorithms that engage in adaptive exploration and combine exploration and exploitation. We consider $Thompson Sampling$ (from hereon $TS$) from this class, which, in a given period, will pull an arm according to the probability that that arm is ``optimal" in the sense of having the highest mean reward \citep{agrawal2012analysis}.
\item ``Naive" algorithms that engage in non-adaptive exploration algorithms and separate exploration and exploitation. We consider $Dynamic$ $\epsilon$-$greedy$ (from hereon $DEG$) from this class, which, in a given period, pulls the arms with the highest posterior mean for $1 - \epsilon$ probability and selects a random arm with $\epsilon$ probability. For our experiments we keep $\epsilon = 0.05$ fixed.
\item Greedy / myopic algorithms that engage in no purposeful exploration and take the best short-sighted action. We consider $DynamicGreedy$ (from hereon $DG$) from this class, which, in a given period, pulls the arm with the highest posterior mean.
\end{enumerate}

It is known that $TS > DEG > DG$ in the standard multi-armed bandit problem of minimizing total regret. The primary question that we want to understand is when, in competition, are the principals incentivized to adopt the ``better" algorithms?

\noindent \textbf{Incumbent} In some of our experiments we modify our model so that one principal enters the market $X$ number of rounds before the other. We refer to the principal that enters before as the ``incumbent" and the principal that enters after $X$ rounds as the ``entrant." The agents that arrive in these $X$ rounds are forced to select the only firm in the market as there is no outside option. We treat $X$ as being an exogenous element of the model and study the consequences for a fixed $X$. We have that both the incumbent and entrant commit to a learning algorithm \textit{before} either firm receives any agent. After the $X$ rounds, both firms still receive the $k$ warm start agents.

\section{Simulation Details}
\label{sim_details}

We evaluate the consequences of our model via simulation.

\textbf{Bandit Priors} We look at bandit instances drawn from three different bandit ``priors" that are different types of learning problems. Recall that we only consider Bernoulli-distributed arms and so we draw the true means of the arms from these priors. For these experiments we fix the number of arms we consider, $K = 10$.
\begin{enumerate}
\item Needle In Haystack - $K-1$ arms with mean 0.5 and 1 arm with 0.7.
\item Uniform - the $K$ arms have means drawn uniformly at random from $[0.25, 0.75]$
\item "Heavy Tail" - the $K$ arms have means drawn from $Beta(\alpha=0.6, \beta = 0.6)$. With this prior it was likely to have means at the ``extremes" or means that were close to 0 as well as means that were close to 1.
\end{enumerate}

\noindent \textbf{Simulation of Competition Game} Unless otherwise noted, all of the reported results utilize the same set of randomly drawn bandit instances and realizations from the prior. Namely, for each bandit prior we draw $N = 1000$ bandit instances. For each of these instances we run simulations of our model for varying values of $k$ and $X$. We take the maximum values of $k$ and $X$ for the simulations we run, $k_{max}$ and $X_{max}$ respectively, and compute a realization table of dimension $(T+k_{max}+X_{max}) \times K$.

This realization table, as well as fixing the random seed for the same bandit instance and realization table across experiments, ensures that differences in algorithm performance are not due to noise in the realizations but due to differences in the algorithms in the different experimental settings. In the competition game we draw from the $T \times K$ portion of the table, so that if two different algorithms picks arm $a$ at time $t$, they get the same $[a, t]$ realization in the table. This setup also ensures that in the warm start period, increasing the warm start from $k$ to $k + 10$ results in the same behavior in the first $k$ rounds.

For the simulations we fix the sliding window size $M = 100$. Low values of $M$ induced too much random noise into the results and we found that increasing $M$ to be larger than $100$ did not make a substantial qualitative difference so we fix this value.

\section{Performance in Isolation}
\label{iso_section}

We evaluate the performance of the different learning algorithms on the bandit priors we consider in isolation. There are two reasons why this is important to consider. First, we want to confirm that the reputation ordering of the algorithms is what we would expect according to the multi-armed bandit literature where $TS > DEG > DG$. Second, we want to better understand what statistics of the instance we can look at in isolation in order to help us predict and understand what to expect in the competition game.

\begin{figure}
\caption{Mean Reputation Trajectories in Isolation}
\includegraphics[scale=0.25]{"figures/nih_iso_mean"}
\label{prelim_means}
\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of a given algorithm on a given instance and then average this value across all the runs. The shaded area display 95\% confidence intervals.}}
\end{figure}

Figure \ref{prelim_means} shows that the mean reputation ordering is as we would expect for the Needle In Haystack prior. The mean reputation ordering that we expect also holds for the other two priors and the results for those can be found in the GitHub appendix.

In the ``competition" game the agents decide between the principals using the \textit{relative} reputation between the two principals. Namely, if principal $A$ has a higher reputation than principal $B$ agents will select principal $A$. A natural question to ask is whether looking at the mean performance is sufficient for understanding performance in the competition game. Figure \ref{rep_dist_nih} displays the reputation distribution at $t = 500$ for the same prior we reported the mean plot for previously. We see that the ``naive" algorithms $DG$ and $DEG$ have a bi-modal reputation distribution whereas $TS$ does not. The intuition for this is that, for Needle In Haystack, $DG$ either finds the best arm or not. If it does, then since it engages in no purposeful exploration it will do better than any algorithm that engages in purposeful exploration over sufficiently many rounds so that the weak law of large numbers comes into play (our $M$ is set sufficiently high to ensure this most of the time). However, if it does not then it will get stuck on a bad arm and lose to $TS$ or $DEG$. In these cases its reputation may be substantially worse but for the competition game the difference between the reputation does not matter but only the binary comparison between them \footnote{This holds for our model and the decision rule of the agents, though the absolute difference may matter if, for instance, considering the SoftMax decision rule in \citet{mansour2017competing}}.

\begin{figure}
\caption{Reputation Distribution}
\includegraphics[scale=0.3]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 200$}}
\end{figure}

This motivates looking at the entire distribution of reputation difference between two algorithms. Figure \ref{ts_dg_rep_diff_nih} shows the distribution of reputation difference between $TS$ and $DG$ across $t$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right. Thus, the mean is not a representative statistic of the entire reputation difference distribution. As an alternative statistic for understanding the results of the competition game we introduce the \textit{relative reputation statistic} which looks at the proportion of simulations in which algorithm $A$ had at least as high of a reputation as algorithm $B$ for a fixed time $t$. This corresponds to running the bandit algorithms in isolation on the same instance and with the same realizations for $t$ rounds and then calculating the fraction of simulations at which an agent would select a principal playing $A$ over a principal playing $B$ at time $t$ \footnote{Though we do not discuss it here, one may be interested in if there is ever a case where, for sufficiently large $t$, we observe that $DEG >DG$ or $TS > DG$ according to the mean reputation but $DG > DEG$ or $DG > TS$ according to the relative reputation proportion. In the GitHub appendix we have results showing that, for Heavy Tail prior with $K=3$ we have that $DEG > DG$ according to the mean reputation but $DG > DEG$ according to the relative reputation proportion.}.

\begin{figure}
\caption{Reputation Difference Distribution}
\includegraphics[scale=0.3]{figures/ts_dg_rep_diff_nih}
\label{ts_dg_rep_diff_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
\end{figure}

Figure \ref{relative_rep_plots} shows the relative reputation plots for $TS$ vs $DG$ on the Uniform and Needle In Haystack prior. For the Uniform prior we see that, in the early rounds, $DG > TS$ for the majority of the simulations but that, eventually, $TS > DG$. The intuition behind this is that, especially since the principals start with no substantive initial information, $TS$ does purposeful exploration in the early rounds in order to acquire information. However, this exploration leads to what we define as a \textit{relative reputation cost}. Namely, exploration leads to lower reputation relative to the myopic alternative. However, eventually the information acquired in the early rounds allows $TS$ to make better decisions and achieve a higher reputation, especially when the instance is ``hard enough" so that $DG$ cannot trivially find the best arm. We define this as the \textit{relative reputation benefit} \footnote{We see similar behavior for the Heavy Tail prior}.

\begin{figure}
\caption{Relative Reputation Plots}
\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
\caption*{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$. The shaded area display 95\% confidence intervals.}
\label{relative_rep_plots}

\end{figure}

Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option.

\section{Competition vs Better Algorithms: Inverted-U}

In this section we document our main results that can be summarized via the empirically documented inverted-U relationship between competition and innovation described in \citet{aghion2005competition}. We consider three separate settings for our model: permanent monopoly, temporary monopoly, and permanent duopoly. In the permanent monopoly and permanent duopoly cases we show that, under certain conditions, firms are incentivized to commit to $DG$ whereas in the temporary monopoly case firms are incentivized to commit to $TS$. We interpret the move from permanent monopoly to temporary monopoly and temporary monopoly to permanent duopoly as increases in competition. We believe this is reasonable and consistent with the empirical results which define increases in competition as increases in a measure of market power (1 - the Lerner Index). In our model we have no prices and no costs and so we interpret increases in ``market power" as decreases in the number of rounds where a firm has to compete for agents.
